Adaptive evolution with aneuploidy and mutation

### 2021-10-26 ms/ms.tex
bit more for prior text

We used growth curves of $2n+1$ and $2n$ in \SI{39}{\celsius} to estimate $1-c+b$. We used the same prior for $w_{2n+1^*}$ and $w_{2n^*}$, assuming that fitness values shouldn't be differ too much. We also tested different priors, but this one (let's call it main prior model) was better in sense of WAIC, prediction plots, estimated parameters. %TODO provide formula of WAIC and the results. To show ppc plots also?. 
\\We tested uninformative uniform prior, $\mathit{Uniform}(1,6)$, for all $w_{2n+1}$, $w_{2n+1^*}$, $w_{2n^*}$ or only for $w_{2n+1^*}$, $w_{2n^*}$. In both scenarios the estimated by the model parameters of fitness were looks too high: for the first scenario the estimated medians were  $w_{2n+1}=2.767$, $w_{2n+1^*}=5.79$, $w_{2n^*}=5.827$ and for the second $w_{2n+1}=1.046$, $w_{2n+1^*}=4.642$, $w_{2n^*}=4.684$. We don't expect to have such high fitness values. Moreover the WAIC for the models with these priors are 0.57 and 1.9 respectively, while WAIC for the main prior model is 0.27. Also for the model with uninformative prior for all fitnesses the fixation time of 2n+1 is about 16 in average, that contradicts unpublished observations that were about 250-300.
\\Also we tried to use additional growth curves. We used growth curves of 2n* (\emph{refined} from \citep{Yona2012}) and 2n+1 in \SI{39}{\celsius} to estimate $w_{2n*}/w_{2n+1}$. The same prior was used for $w_{2n*}/w_{2n+1*}$. This model not perfect well with WAIC 2.8.
\\Another try was to use growth curves of $2n+1$ and $2n$ in \SI{30}{\celsius} to estimate $1-c$. This estimation assumes that the cost of aneuploidy is the same in \SI{39}{\celsius} and \SI{30}{\celsius}; this might be incorrect, but we only assumed this to generate a prior distribution for the fitness values. The prior for $b$ was taken the same as for $c$. This model didn't work at all: we didn't receive any parameter that the likelihood with it is greater than zero. %TODO it make sense because bla bla?

Because \texttt{Curveball} uses a maximum-likelihood approach to estimate model parameters, we were able to estimate a distribution of relative fitness values by sampling from a truncated multivariate normal distribution defined by the maximum-likelihood covariance matrix. Thus, we sampled 10,000 values for appropriate fitness, which we used as prior distributions. 
See Figures~\ref{fig:growth-curves-30deg} and~\ref{fig:growth-curves-39deg}.
%TODO ILIA please report here the point estimate and 95% CI for the three distributions, I only have for 1-c+b: 0.9863 (95% CI 0.9790 - 0.9939).
%TODO ILIA describe the priors for the mutation and aneuploidy rate.
% TODO also describe the prior for the model without aneuploidy - the mutation rate is wider here.

>>>>

We used growth curves of $2n+1$ and $2n$ in \SI{39}{\celsius} to estimate  $w_{2n+1}$. We used the same prior for $w_{2n+1^*}$ and $w_{2n^*}$, assuming that fitness values shouldn't be differ too much. We also tested different priors, but this one (let's call it main prior) was better in sense of WAIC, prediction plots, estimated parameters. %TODO provide formula of WAIC. Table? To show ppc plots also?. 
\\We tested uninformative uniform prior, $\mathit{Uniform}(1,6)$, for all $w_{2n+1}$, $w_{2n+1^*}$, $w_{2n^*}$ and only for $w_{2n+1^*}$, $w_{2n^*}$, leaving $w_{2n+1}$ prior as before . In both scenarios the estimated by the model parameters of fitness were looks too high: for the first scenario the estimated medians were  $w_{2n+1}=2.767$, $w_{2n+1^*}=5.79$, $w_{2n^*}=5.827$ and for the second $w_{2n+1}=1.046$, $w_{2n+1^*}=4.642$, $w_{2n^*}=4.684$. We don't expect to have such high fitness values. Moreover the WAIC for the models with these priors are 0.57 and 1.9 respectively, while WAIC for the main prior model is 0.27. Also for the model with uninformative prior for all $w_{2n+1}$, $w_{2n+1^*}$, $w_{2n^*}$ the fixation time of 2n+1 is about 16 in average, that contradicts unpublished observations that were about 250-300.
\\Also we tried to use additional growth curves. We used growth curves of 2n* (\emph{refined} from \citet{Yona2012}) and 2n+1 in \SI{39}{\celsius} to estimate $w_{2n*}/w_{2n+1}$. The same prior was used for $w_{2n*}/w_{2n+1*}$. This model didn't perform well, resulted in WAIC 2.8.
\\Another try was to use growth curves of $2n+1$ and $2n$ in \SI{30}{\celsius} to estimate $1-c$. This estimation assumes that the cost of aneuploidy is the same in \SI{39}{\celsius} and \SI{30}{\celsius}; this might be incorrect, but we only assumed this to generate a prior distribution for the fitness values. The prior for $b$ was taken the same as for $c$. This model didn't work at all: we didn't receive any parameter that the likelihood with it is greater than zero. %TODO it make sense because bla bla?

Because \texttt{Curveball} uses a maximum-likelihood approach to estimate model parameters, we were able to estimate a distribution of relative fitness values by sampling from a truncated multivariate normal distribution defined by the maximum-likelihood covariance matrix. Thus, we sampled 10,000 values for appropriate fitness, which we used as prior distributions. 
See Figures~\ref{fig:growth-curves-30deg} and~\ref{fig:growth-curves-39deg}.

For the mutation rate $\mu$ and aneuploidy rate $\delta$ we used uniform priors, $\mathit{Uniform}(1e-8,1e-5)$ and $\mathit{Uniform}(1e-6,1e-3)$ respectively. We have tried also uniform priors with lower than $1e-8$ bound for $\mu$, but anyway the posterior ends with higher values. 
For the single locus model without aneuploidy, for the mutation rate wider uniform prior was used - $\mathit{Uniform}(1e-10,1e-5)$. 
%TODO ILIA please report here the point estimate and 95% CI for the three distributions, I only have for 1-c+b: 0.9863 (95% CI 0.9790 - 0.9939).
%TODO ILIA describe the priors for the mutation and aneuploidy rate.
% TODO also describe the prior for the model without aneuploidy - the mutation rate is wider here.
